---
title: "Exercise: Matching"
output: pdf_document
fontsize: 16pt
---

```{r setup, echo=F,warning=F, error=F, message=F}
knitr::opts_chunk$set(echo = F, warning=F, error=F, message=F, fig.height=3, fig.width=3, include=F)
library(tidyverse)
library(stargazer)
library(MatchIt)
```

Let's simulate some fake data and see when we are able to recover the correct treatment effect using matching methods. 

1. First, let's generate some confounder variables for 100 people. 
(a) The variable 'age' should be drawn randomly from the normal distribution with mean 40 and standard deviation 7.
(b) The variable 'gender' should be drawn randomly from the binomial distribution with a 0.5 probability of being male or female.
(c) The variable 'income' should be drawn randomly from the normal distirbution with mean 500 and standard deviation 50.
(d) The variable 'education' should be randomly drawn from one of four numerical categories with equal probability: 0 (None), 1 (Primary), 2 (Secondary), 3 (Tertiary)

```{r}
set.seed(54321)
N <- 100
d <- tibble(age=rnorm(N,40,7),
            gender=rbinom(N,1,0.5),
            income=rnorm(N,500,50),
            education=sample(c(0,1,2,3),N,
                             prob=c(0.25,0.25,0.25,0.25), replace=T))
```

2. Our outcome is going to be attitudes to redistribution. Use the expressions below to simulate potential outcomes, with a treatment effect of 5.

$$y_0=N(10,2)+5*age - 100*gender + income - education*30$$ 

$$y_1=y_0+5$$

```{r}
set.seed(54321)
d <- d %>% mutate(y_0=rnorm(N,10,2) + 5*age - 100*gender + income - education*30,
                  y_1=y_0+5)
```

3. Treatment $D$ is receiving a government social program, but treatment is **not** randomly  assigned in any way. Instead, treatment depends on age, gender, income and education. Binary (1/0) treatment is determined by the following expression:

$$D=\begin{cases}
1 \text{ if } 2*gender + \frac{age}{2} + \frac{income}{50} + education*2>34 \\ 
0 \text{ else }
\end{cases}$$

```{r}
d <- d %>% mutate(D=case_when(2*gender+age/2+income/50+education*2>34~1,
                              T~0))

summary(2*d$gender + d$age/2 + d$income/50 + d$education*2)
```

4. Calculate observed outcomes based on potential outcomes and treatment.

```{r}
d <- d %>% mutate(y_obs=case_when(D==0~y_0,
                                  D==1~y_1))
```

5. As always, as a benchmark, let's run the 'naive' regression of the outcome on the treatment with no controls. Why is the result different from our assumed treatment effect?

```{r}
d %>% lm(y_obs ~ D, data=.) %>% summary()
```

6. Our first task is to try and 'match' one treated unit with one control unit so that the *only* thing that is different about them is their treatment status. Take the first treated unit in your dataset. What are its values of gender, age, income and education? Manually, by trial-and-error (not using any package or pre-prepared function), identify the most similar *control* unit.

```{r}
treated_unit <- d %>% filter(D==1) %>% slice(1)
control_units <- d %>% filter(D==0 & gender==1 & education==3)
control_unit <- control_units %>% filter(age<36 & age>32 & income>450 & income<570)
```

7. Compare the outcome between your matched treated unit and control unit. Is this consistent with our assumed treatment effect? Why/why not?

```{r}
treated_unit$y_obs - control_unit$y_obs
```

8. Matching repeats this process for multiple units and then finds the average difference in outcomes between the treated and control units. Use the _matchit_ package to conduct 'exact' matching on your dataset for the two discrete confounder variables: gender and education. What is the result of the matching procedure?

```{r}
matched_data_Q8 <- matchit(D ~ gender + education, data=d, method="exact")
```

9. Use _match.data_ to extract the matched dataset and calculate the average difference in means between the treated and control groups. How does the result compare to the naive regression in Q5?

```{r}
match.data(matched_data_Q8) %>% group_by(D) %>%
  summarize(y_obs=mean(y_obs,na.rm=T)) %>%
  arrange(-D) %>%
  mutate(diff_y_obs=y_obs-lead(y_obs))
```

10. Matching *ONLY* makes a difference if we throw away some data - the data for which we cannot find good matches. The more data we throw away, the better matched/balanced is our remaining data.

```{r}

```

