---
title: "FLS 6415: Replication 3 - Natural Experiments"
#author: "Jonathan"
date: "April 2020"
output:
  pdf_document: default
  html_document: default
---

To be submitted (code + answers) by midnight, Wednesday 8th April.

First read the paper by De La O (2013) on the class website. 

The replication data is in the file *DelaO.csv*, and the important variables are described below. Each row of this dataset is one electoral precinct, some of which are considered treated because they received *Progresa*. 

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(haven)
library(tidyverse)
library(knitr)
library(broom)
library(stargazer)
library(estimatr)
library(texreg)

opts_chunk$set(echo=F, warning=F, message=F, include=F)
```

```{r variables, echo=F, include=F, message=FALSE, warning=FALSE}
d <- read_dta("DeLaO_AJPS2013_rep_file.dta")

d %>% write_csv("DeLaO.csv")
d <- read_csv("DeLaO.csv")
```

```{r, echo=F, warning=FALSE, message=FALSE, include=T}
vars <- data.frame(names(d)[c(7,8,9,5,4,6,13,20,22,24,26,16,15,17,18,19,21,23,25)],c("Whether the precinct received Progresa","Number of Treated Villages in Precinct","Number of Control Villages in Precinct","Poverty in 1995","Population in 1994/5","Population Eligible","Number of Villages in Precinct","Turnout % in 1994","PRI vote share 1994","PAN vote share 1994","PRD vote share 1994","Total Number of Votes in 1994","Number of PRI votes in 1994","Number of PAN votes in 1994","Number of PRD votes in 1994","Turnout % in 2000","PRI Vote Share in 2000","PAN Vote Share in 2000","PRD Vote Share in 2000"))
colnames(vars) <- c("Variable","Description")

kable(vars,caption="Key Variables in De La O (2013)")
```


**1. First, what is treatment in this study? What is control? What is the outcome being measured?**

**2. To help assess the balance between treatment and control units, reproduce Table 2 in De La O (2013) (Don't worry about the standard errors in brackets in the 'Difference' column for now).**

```{r, warning=FALSE}
tab2_vars <- names(d)[c(5,4,6,13,10,11,20,22,24,26)]

Q1 <- d %>% group_by(treatment) %>% 
  summarise_at(tab2_vars,mean,na.rm=TRUE) %>% 
  gather(key,value,-treatment) %>% 
  spread(treatment,value) %>% 
  rename(Variable="key",Early=`1`,Late=`0`) %>%
  mutate(Difference=Late-Early)

kable(Q1,caption="Q2: Balance in the Data", digits=3)

```

**3. Is the balance shown in this table (Table 2 in De La O) a necessary condition for causal inference? Is it a sufficient condition for causal inference?**

**4. The main analysis in De La O is conducted on a subset of the full dataset. Filter the data so that only precincts that have either one treatment village (`numerotreated`) or one control village (`numerocontrol`) inside them are included in your new dataset. What percentage of the original precincts are included in the new dataset?**

```{r}
subset_d <- d %>% filter(numerotreated==1|numerocontrol==1)

pct_data <- round(dim(subset_d)[1]/dim(d)[1]*100,1)

```

**5. One of De La O's conclusions is that treatment (receiving Progresa) boosts turnout. Conduct a simple difference-in-means t-test on the filtered dataset from Q4 to assess this claim. What is the estimated difference-in-means and how statistically significant is the result? **

```{r}
subset_d %>% t.test(t2000~treatment, data=.) %>% 
  tidy() %>%
  kable(caption="Q5: Difference-in-Means")

```

**6. De La O's analysis of turnout is in the upper panel of Table 3, where she runs a regression, adding some controls. (We are going to focus on the 'ITT' estimates, we will talk about the 'IV' estimates next week). Replicate this turnout regression. The controls (listed under De La O's Table 3) are `avgpoverty,pobtot1994,votos_totales1994,pri1994,pan1994,prd1994` and there is a fixed effect for the *villages* variable. (Try to include the robust standard errors, but no problem if you cannot). Interpret the results. **

```{r, results='asis'}
reg1 <- lm_robust(t2000~treatment + avgpoverty + pobtot1994 + votos_totales1994 + pri1994 + pan1994 + prd1994 + factor(villages),data=subset_d, se_type="HC1")
texreg(reg1, include.ci=F, digits=3, caption="Q6: With Fixed Effects")
```


**7. Now run the same regression but exclude the number-of-villages fixed effects (keep the other controls). How does this change the comparisons we are making between treated and control villages? How do the results change?**

```{r, results='asis'}
reg2 <- lm_robust(t2000~treatment + avgpoverty + pobtot1994 + votos_totales1994 + pri1994 + pan1994 + prd1994,data=subset_d, se_type="HC1")
texreg(reg2, include.ci=F, digits=3, caption="Q7: Without village Fixed Effects")
```

**8. Replicate all four columns of the upper panel of Table 3 in De La O (2013). Interpret the results. **

```{r, results='asis'}
dvs <- c("t2000","pri2000s","pan2000s","prd2000s")

models <- dvs %>% map(~lm_robust(formula(paste(.x, "~treatment + avgpoverty + pobtot1994 + votos_totales1994 + pri1994 + pan1994 + prd1994 + factor(villages)")),
                                 data=subset_d, 
                                 se_type="HC1"))
texreg(models, include.ci=F, digits=3, caption="Q8: De La O Table 3")
```

**9. Now let's look at some critiques of the paper. Normally, we measure turnout percentages and vote shares as being naturally bounded between 0 and 100% (or 0 and 1). Other numbers don't make sense. Use a boxplot or similar graph to assess the distribution of values on the four dependent variables. What do you find?**

```{r, fig.cap="Non-feasible values in the Dependent Variables"}
subset_d %>% select(dvs) %>% 
  gather(key="Variable",value="Value") %>% 
  ggplot() +
  geom_boxplot(aes(x=Variable, y=Value)) +
  theme_classic()

```

**10. As a 'quick fix' replace all the unrealistic values above 100% (1) with `NA` for all the turnout percentage and vote share dependent variables. Re-run your regressions from question 8. Do your conclusions change? Why might this be?**

```{r, echo=TRUE,results='asis'}
corrected_d <- subset_d %>% mutate_at(dvs, function(x) {ifelse(x>1, NA,x)})

models2 <- dvs %>% map(~lm_robust(formula(paste(.x, "~treatment + avgpoverty + pobtot1994 + votos_totales1994 + pri1994 + pan1994 + prd1994 + factor(villages)")),
                                 data=corrected_d, 
                                 se_type="HC1"))
texreg(models2, include.ci=F, digits=3, caption="Q10: Feasible values for Dependent Variables")

```

**11. Next, examine the control variable for population in 1994 (`pobtot1994`). Use a graph or other method to identify any extreme outliers. Extreme values of control variables are not a problem if they are balanced across treatement and control groups. But are they in this case? Identify whether the extreme outliers are in the control or treatment group.**

```{r, fig.cap="Q11: Outliers in Population Data"}
subset_d %>% 
  select(treatment, pobtot1994) %>% 
  ggplot() +
  geom_boxplot(aes(y=pobtot1994, group=treatment)) +
  theme_classic()
```

**12. Remove the extreme outliers you identified in Q11 from the dataset (the dataset before you removed the infeasible values of the dependent variables). Re-run your regressions. Do your conclusions change? Why might this be?**

```{r, echo=TRUE, results='asis'}
popn_corrected_d <- subset_d %>% mutate(rank_popn=rank(-pobtot1994))  %>% 
  filter(rank_popn>2)

models3 <- dvs %>% map(~lm_robust(formula(paste(.x, "~treatment + avgpoverty + pobtot1994 + votos_totales1994 + pri1994 + pan1994 + prd1994 + factor(villages)")),
                                 data=popn_corrected_d, 
                                 se_type="HC1"))
texreg(models3, include.ci=F, digits=3, caption="Q12: Removing Population Outliers")

```

**13. One more issue. The controls for the regressions you have conducted so far are the *absolute number* of votes for turnout, PRI, PAN and the PRD. But for the dependent variable, De La O is using the *percentage vote share of the population*. Arguably it might be more consistent to use the same measurement approach on both the left and right-hand sides of the regression. Try implementing the regressions using the controls `t1994, pri1994s, pan1994s, prd1994s` in place of `votos_totales1994, pri1994, pan1994, prd1994`. Ignore the other corrections you made in previous questions. Does this change your conclusions? Why might this be?**

```{r, echo=TRUE, results='asis'}
models4 <- dvs %>% map(~lm_robust(formula(paste(.x, "~treatment + avgpoverty + pobtot1994 + t1994 + pri1994s + pan1994s + prd1994s + factor(villages)")),
                                 data=popn_corrected_d, 
                                 se_type="HC1"))
texreg(models4, include.ci=F, digits=3, caption="Q13: Using Percentage Controls")
```
