---
title: "Exercise: Comparative Case Studies & Process Tracing"
output: pdf_document
fontsize: 16pt
---

```{r setup, echo=F,warning=F, error=F, message=F}
knitr::opts_chunk$set(echo = F, warning=F, error=F, message=F, fig.height=3, fig.width=3, include=F)
library(tidyverse)
library(stargazer)
library(MatchIt)
library(optmatch)
library(rgenoud)
library(cem)
library(zeligverse)
```

Let's simulate some fake data and see whether we are able to recover the correct treatment effect using comparative case study methods. Our research question is whether development causes democratization.

1. First, let's generate some confounder variables for 10 countries (we have only cross-section data).   
(a) The variable 'Inequality' should be drawn randomly with equal probability from the values "high" and "low". 
(b) The variable 'Economic Growth' should be drawn randomly with equal probability from the values "high" and "low".
(c) The variable 'Continent' should be randomly drawn with equal probability from four possibilities: Europe, Americas, Asia, and Africa.

```{r}
set.seed(54321)
N <- 10
d <- tibble(inequality=sample(c("High","Low"),N,prob=c(0.5,0.5),replace=T),
            growth=sample(c("High","Low"),N,prob=c(0.5,0.5),replace=T),
            country=sample(c("Europe","Americas","Asia","Africa"),N,
                             prob=c(0.25,0.25,0.25,0.25), replace=T))
```

2. Our outcome is going to be a binary variable for whether the country is a democracy or not. Use the expressions below to simulate potential outcomes, with a treatment effect which boosts the chances of democratization by 50\% (so our treatment effect is 0.5).

$$y_0= \begin{cases}
1\text{ if } N(2,2) + 2*inequality - growth + Europe>0\\ 
0 \text{ else}
\end{cases}$$ 

$$y_1=\begin{cases}
1\text{ if } y_0=1\\
1\text{ if } y_0=0 \text{ } \& \text{ } Binom(0.5)=1\\
0 \text{ else}
\end{cases}$$

```{r}
set.seed(54001)
d <- d %>% mutate(Europe=ifelse(country=="Europe",1,0),
  y_0=case_when(rnorm(N,2,2) + 2*High_inequality - High_growth + 2*Country_Europe > 0~1,
                                T~0),
  rnd_y_1 = rbinom(N,1,0.5),
                  y_1=case_when(y_0==1~1,
                                y_0==0 & rnd_y_1==1~1,
                                T~0))
```

3. Treatment $D$ is the level of income of the country. It depends on the confounding variables and we will simplify to High versus Low income based on whether the country has an income over USD10,000. Imagine we know the treatment assignment mechanism so that binary (1/0) treatment is determined by the following expression:

$$D=\begin{cases}
1 \text{ if } inequality>0.5 &  - 1000*inequality - growth*500 > 10000 \\ 
0 \text{ else }
\end{cases}$$

```{r}
set.seed(54001)
d <- d %>% mutate(D=case_when(2000*Europe-1000*inequality-growth*500 + rnorm(N,10000,2000)>10000 ~ 1,
                              T~0))
```

4. Calculate observed outcomes based on potential outcomes and treatment.

```{r}
d <- d %>% mutate(y_obs=case_when(D==0~y_0,
                                  D==1~y_1))
```

5. As always, as a benchmark, let's run the 'naive' regression of the outcome on the treatment with no controls. 

```{r}
d %>% lm(y_obs ~ D, data=.) %>% summary()
```

6. Now we will implement a comparative case study methodology for just two units. Select a treated (high income) unit that is a democracy. What are the values of the three confounding variables; inequality, growth and country? 

```{r}

```

7. From all of the control (low income) units that are NOT democracies, pick one at random. Is it possible that inequality explains why your treated unit is a 
