---
title: "Exercise: Comparative Case Studies & Process Tracing"
output: pdf_document
fontsize: 16pt
---

```{r setup, echo=F,warning=F, error=F, message=F}
knitr::opts_chunk$set(echo = F, warning=F, error=F, message=F, fig.height=3, fig.width=3, include=F)
library(tidyverse)
library(stargazer)
library(MatchIt)
library(optmatch)
library(rgenoud)
library(cem)
library(zeligverse)
```

Let's simulate some fake data and see whether we are able to recover the correct treatment effect using comparative case study methods. Our research question is whether development causes democratization.

1. First, let's generate some confounder variables for 100 countries (we have only cross-section data).   
(a) The variable 'Inequality' should be drawn randomly from the normal distribution with mean 0.5 and standard deviation 0.2. 
(b) The variable 'Economic Growth' should be drawn randomly from the normal distribution with mean 3 and standard deviation 2. 
(c) The variable 'Continent' should be randomly drawn with equal probability from four possibilities: Europe, Americas, Asia, and Africa.

```{r}
#rm(list=ls())
set.seed(54321)
N <- 100
d <- tibble(inequality=rnorm(N,0.5,0.2),
            growth=rnorm(N,3,2),
            country=sample(c("Europe","Americas","Asia","Africa"),N,
                             prob=c(0.25,0.25,0.25,0.25), replace=T))
```

2. The relationships we will assume are based on existing theory and evidence:
  * European countries are more likely to be high-income and democracies
  * 

2. Our outcome is going to be a binary variable for whether the country is a democracy or not. Use the expressions below to simulate potential outcomes, with a treatment effect which boosts the chances of democratization by 50\% (so our treatment effect is 0.5). (`Europe` is an indicator for whether the country is in Europe), `Binom(0.5)` is a random binary variable with probability 0.5 of being 1).

$$y_0= \begin{cases}
1\text{ if } (2*inequality - \frac{growth}{10} + Europe) >1\\ 
0 \text{ else}
\end{cases}$$ 

$$y_1=\begin{cases}
1\text{ if } y_0=1\\
1\text{ if } y_0=0 \text{ } \& \text{ } Binom(0.5)=1\\
0 \text{ else}
\end{cases}$$

```{r}
set.seed(54001)
d <- d %>% mutate(Europe=ifelse(country=="Europe",1,0),
  y_0=case_when(2*inequality - growth/10 + Europe>1~1,
                                T~0),
  rnd_y_1 = rbinom(N,1,0.5),
                  y_1=case_when(y_0==1~1,
                                y_0==0 & rnd_y_1==1~1,
                                T~0))

#summary(2*d$inequality - d$growth/10 + 0.25)
```

3. Treatment $D$ is the level of income of the country. It depends on the confounding variables and we will simplify to High versus Low income based on whether the country has an income over USD10,000. Imagine we know the treatment assignment mechanism so that binary (1/0) treatment is determined by the following expression:

$$D=\begin{cases}
1 \text{ if } (N(10000,2000) + 2000*Europe - 1000*inequality - growth*200) > 10000 \\ 
0 \text{ else }
\end{cases}$$

```{r}
set.seed(54001)
d <- d %>% mutate(D=case_when(2000*Europe-1000*inequality-growth*200 + rnorm(N,10000,2000)>10000 ~ 1,
                              T~0))
```

4. Calculate observed outcomes based on potential outcomes and treatment.

```{r}
d <- d %>% mutate(y_obs=case_when(D==0~y_0,
                                  D==1~y_1))
```

5. Draw the causal diagram that represents the relationships in the simulated data. (Make sure you indicate the direction of the effect).

6. 

5. As always, as a benchmark, let's run the 'naive' regression of the outcome on the treatment with no controls. 

```{r}
d %>% lm(y_obs ~ D, data=.) %>% summary()
```

6. Now we will implement a comparative case study methodology for just two units. Select one treated (**high income**) unit that is a **democracy**. What are the values of the three confounding variables, inequality, growth and contintent? 

```{r}
set.seed(123)
d_treat <- d %>% filter(D==1 & y_obs==1) %>% sample_n(1)
ineq_treated <- d_treat %>% pull(inequality)
growth_treated <- d_treat %>% pull(growth)
country_treated <- d_treat %>% pull(country)
```

7. From all of the control (low income) units that are NOT democracies, pick one at random. 
(a) Is it *possible* that inequality explains why your treated unit is a democracy and the control unit is not?
(b) Is it *possible* that economic growth explains why your treated unit is a democracy and the control unit is not?
(c) Is it *possible* that being European explains why your treated unit is a democracy and the control unit is not?
(d) Can we conclude anything about the role of income level (treatment) on democracy?

```{r}
set.seed(123)
d %>% filter(D==0 & y_obs==0) %>% sample_n(1)
```

8. Identify one control unit that has values of the confounders that ensure you can answer "No" to questions (a)-(c) in Question 7. (Note this may not mean the closest unit but the unit with values of the variables in the 'right' direction) It may not be possible - if not, try picking a different treated unit and then search for an appropriate control unit.

```{r}
d %>% filter(D==0 & y_obs==0 & country==country_treated & growth<growth_treated & inequality>ineq_treated)
```

9. Once you succeed in finding an appropriate pairing of treated and control units, what can you conclude about the role of development in explaining democracy?

10. How \textbf{representative} are your treated and control cases? Use a table or graph or numerical comparison to compare them to the overall distribution of countries on the confounding variables.

11. Return to your selected treated case and ignore the control cases. If we were to conduct process tracing on our single treated case to generate additional evidence that it was development that caused democracy in this case, what observable implications of this theory/mechanism would you look for? 

12. A competing theory of democratization is the role of inequality in generating a stronger demand for redistribution, which is expressed through protests by the left and is ultimately only possible under a democracy. What observable implications can you think of that might help us \textit{exclude} the mechanism of inequality as an explanantion of our treated case using process tracing? 

13. Comparing the observable implications you have suggested in Q11 and Q12, which are most likely to help us distinguish the influence of development versus inequality?

