---
title: "Exercise: Discontinuities"
output: pdf_document
fontsize: 16pt
---

```{r setup, echo=F,warning=F, error=F, message=F}
knitr::opts_chunk$set(echo = F, warning=F, error=F, message=F, fig.height=3, fig.width=3, include=F)
library(tidyverse)
library(stargazer)
library(broom)
```

This week we conduct the same simulation as usual, but using a regression discontinuity to estimate the effect. 

1. First, let's generate an 'income' variable for 10,000 people. The data should be drawn randomly from the normal distribution with mean 500 and standard deviation 50.

```{r}
set.seed(54321)
N <- 50000
d <- tibble(income=rnorm(N,500,50))
```

2. Now let's simulate potential outcomes (let's say the outcome is 'attitude to redistribution') for each person that depend on their income. Assume:

$$y_0=N(10,2)+\frac{income}{100}$$ 

$$y_1=y_0+2$$

So there is a constant treatment effect of 2.

```{r}
d <- d %>% mutate(y_0=rnorm(N,10,2) + income/100,
                  y_1=y_0+2)
```

3. Actual treatment is not random but deterministic: Imagine it is a poverty-relief program that you receive only if your income is less than 450.

```{r}
d <- d %>% mutate(D=ifelse(income<500,1,0))
```

4. Now calculate the observed outcome based on the potential outcomes and actual treatment status.

```{r}
d <- d %>% mutate(y_obs=case_when(D==0~y_0,
                                  D==1~y_1))
```

5. Now let's calculate the 'naive' Average Treatment Effect by running a simple OLS regression of the observed outcomes on treatment. What is your estimate of the average treatment effect? How does this compare to the treatment effect we specified earlier?

```{r, results='asis'}
d %>% lm(y_obs ~ D, data=.) %>% stargazer(single.row=T, header=F, title="Q7")
```

6. Now apply the 'full bandwidth' regression discontinuity method. This just means controlling for the running variable, which in this case is income. Interpret the results. How do they compare to the treatment effect we specified?

```{r}
d %>% lm(y_obs ~ D + income, data=.) %>% summary()
```

7. Let's try and make a regression discontinuity plot to understand what's going on better. First, plot the running variable (which one is that?) against the observable outcome variable. What can you see in the graph?

```{r}
d %>% ggplot() + 
  geom_point(aes(x=income, y=y_obs)) +
  theme_classic()
```

8. The graph in Q7 is messy so most regression discontinuity plots 'bin' the data into smaller groups to more easily see the pattern. Bin the income data into 20 groups and plot this against the average observed outcome in each bin. 

```{r}
d %>% mutate(income_binned=cut(income,20)) %>%
  group_by(income_binned) %>%
  summarize(y_obs_mean=mean(y_obs,na.rm=T)) %>%
  ggplot() +
  geom_point(aes(x=income_binned,y=y_obs_mean)) +
  geom_vline(xintercept=500, lty=2, col="red") +
  theme_classic() +
  ylim(12,17)
```

9. An alternative way of estimating the treatment effect is to perform a simple difference-in-means between the values just to the left and just to the right of the cutoff. Using the data between 480 and 520 on the income scale, perform a difference-in-means test for the efffect of treatment on the outcome. How does this compare to your regression discontinuity estimate?

```{r}
d %>% filter(income<520 & income>480) %>%
  t.test(y_obs~D, data=.) %>%
  tidy()
```

10. A third method of performing a regression discontinuity is to perform the regression approach only on a narrow bandwidth close to the cutoff. Subset the data again to between 480 and 520 on the income scale and perform a regression discontinuity analysis. How do the results compare to your full-bandwidth regression discontinuity in Q6?

```{r}
d %>% filter(income<520 & income>480) %>% 
  lm(y_obs ~ D + income, data=.)
```

11. 

```{r}
d %>% lm(y_obs ~ D + income + income^2, data=.) %>% summary()
```



11. Compare the confidence intervals on the treatment effect from regression discontinuities with the following bandwidths:

```{r}
bandwidths <- tibble(lower=c(400,420,440,460,480),
       upper=c(600,580,560,540,520))

bandwidths %>% map2(filter,d)


d %>% filter(income<520 & income>480) %>% 
  lm(y_obs ~ D + income, data=.)
```



9. If you can, add a vertical line where the treatment cutoff line is. Then try to add one regression line to the left of the cutoff and one regression line to the right of the cutoff.

```{r}
d %>% mutate(income_binned=cut(income,20)) %>%
  mutate(income_midpoint=)
  group_by(income_binned) %>%
  summarize(y_obs_mean=mean(y_obs,na.rm=T)) %>%
  ggplot() +
  geom_point(aes(x=income_binned,y=y_obs_mean)) +
  geom_vline(xintercept=500, lty=2, col="red") +
  geom_smooth(aes(x=income_binned, y=y_obs_mean)) +
  theme_classic() +
  ylim(12,17)
```



7. A
