---
title: "Exercise: Understanding Potential Outcomes"
output: pdf_document
fontsize: 16pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, include=F)
library(tidyverse)
```

1. Generate data on a population of 1,000 people. Specifically, create a variable (a vector) that randomly assigns these people to be male or female (50:50). *Hint: In R, try $rbinom$ and in Stata, try $rbinomial$.*
```{r}
N <- 1000
x <- rbinom(N,1,0.5)
```

2. Now we are going to simulate the potential outcomes - a measure of attitudes - *under control* ($y_0$) for our population. Create another variable of random normally-distributed values with mean of 5 and standard deviation of 1. 

```{r}
y0 <- rnorm(N,5,1)
```

3. One problem with observational data is that potential outcomes are often correlated with other variables such as gender. Adjust your value of $y_0$ to add 1 (one) for all units who are male.
```{r}
y0 <- y0 + x
```

4. Now simulate potential outcomes *under treatment* ($y_1$) for all units. Define a *constant* treatment effect of $c=2$  and create another vector $y_1=y_0+c$.
```{r}
c <-2
y1 <- y0 + c
```

5. To compare our two sets of potential outcomes, plot two density charts on the same figure - one for $y_0$ and one for $y_1$.
```{r}
data <- tibble(x,y0,y1)

data %>% ggplot() +
  geom_density(aes(x=y0), col="blue") +
  geom_density(aes(x=y1),col="dark green") +
  theme_classic()
```

6. Next, let us assume a specific **Treatment Assignment Mechanism** where men are more likely than women to receive treatment. First, create a temporary 'latent' variable which consists of two components added together: (i) $0.5*x$ (i.e. half the value of the gender variable), and (ii) a random uniform value between zero and one. Finally, create a new vector $D$ which is equal to one when the latent variable is larger than 0.75, and zero otherwise.
```{r}
data <- data %>% mutate(rnd=runif(N,0,1),
                        D=ifelse(0.5*x+rnd>0.75,1,0))
```

7. To show that gender and treatment - maybe we can think of treatment as low income and high income - are related, calculate the correlation between $x$ and $D$.
```{r}
cor(data$x,data$D)
```

8. What is the average of the *real* indvidual treatment effects based on the potential outcomes, $E(y_1-y_0)$? 
```{r}
Actual_causal_effect <- data %>% 
  summarize(Actual_ATE=mean(y1-y0))
```

9. The Fundamental Problem of Causal Inference is that we *cannot* calculate (8.) above. Instead, we only observe one value: $y_{obs}$. Create a new variable $y_{obs}$ which equals $y_1$ if $D=1$ but which equals $y_0$ if $D=0$.
```{r}
data <- data %>% mutate(y_obs=case_when(D==1~y1,
                                        D==0~y0))
```

10. Based on the observable data, run the basic regression of treatment ($D$) on observable outcomes ($y_{obs}$). Interpret the result. Is this an accurate estimate of the treatment effect that we assumed at the start?
```{r}
data %>% lm(y_obs~D,data=.) %>% summary()
```

11. Re-run all your code above but this time with $c=0$ so we are assuming **NO** treatment effect. Run the regression in (10.) again - what is the result?

```{r}
data_no_effect <- tibble(x=rbinom(N,1,0.5),
                         y0=x+rnorm(N,5,1),
                         y1=y0+0,
                         rnd=runif(N,0,1),
                         D=ifelse(0.5*x+rnd>0.75,1,0)) %>% 
  mutate(y_obs=case_when(D==1~y1,
                         D==0~y0))

data_no_effect %>% ggplot() +
  geom_density(aes(x=y0), col="blue") +
  geom_density(aes(x=y1),col="dark green") +
  theme_classic()

data_no_effect %>% lm(y_obs~D,data=.) %>% summary()
```

12. To see why, let's plot two density charts on the same figure - one for the distribution of observable $y_{obs}$ for the treated group ($y_{obs}|D==1$) and one for the distribution of observable $y_{obs}$ for the control group ($y_{obs}|D==0$).
```{r}
data_no_effect %>% ggplot() +
  geom_density(data=data_no_effect %>% filter(D==0),aes(x=y_obs), col="blue") +
  geom_density(data=data_no_effect %>% filter(D==1),aes(x=y_obs),col="dark green") +
  theme_classic()
```

13. Run your code again for $c=0$, but this time assume a larger population of $N=1,000,000$. Does that solve the problem?

```{r}
N <- 1000000
data_large_N <- tibble(x=rbinom(N,1,0.5),
                         y0=x+rnorm(N,5,1),
                         y1=y0+0,
                         rnd=runif(N,0,1),
                         D=ifelse(0.5*x+rnd>0.75,1,0)) %>% 
  mutate(y_obs=case_when(D==1~y1,
                         D==0~y0))

data_large_N %>% ggplot() +
  geom_density(aes(x=y0), col="blue") +
  geom_density(aes(x=y1),col="dark green") +
  theme_classic()

data_large_N %>% lm(y_obs~D,data=.) %>% summary()
```

14. For $c=0$, run the regression of treatment on observable outcomes, but this time controlling for gender.
```{r}
data_no_effect %>% lm(y_obs~D + x,data=.) %>% summary()
```

